---
title: "Final Project"
output: html_document
---

## Run to get started

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

# Run to load in results of the data from FB Reference
```{r}
Prem <- read_csv("https://raw.githubusercontent.com/ashwinstratton/Forecasting-Final-Datasets/main/PremData.csv")
```
## Run to see the data that you just loaded
```{r}
Prem
```

## This histogram provides us the distribution of the data on a visual perspective (Poisson)
```{r}
ggplot(data = Prem, aes(x=Gls)) +
  geom_bar(stat="count")
```

Model : Poisson

## Model 2 is what we were happiest with, in terms of statistically significant variables

```{r}
PremScorersGLM <- glm(Gls ~ PrgC...23 + Sh + GCA + TO...78 + PKatt...16 , family = "poisson", data=Prem) 
#Variables - Progressive Carries, Shots, Goal Creating Actions, Take-Ons, Penalty Kick Attempts
summary(PremScorersGLM)
```

```{r}
PremScorersGLM2 <- glm(Gls ~ PrgC...23 + Sh + SoTPercent + SCA + Dist + Prem$`90s` , family = "poisson", data=Prem) 
#Variables - Progressive Carries, Shots, Shots on Targer Percentage, Shot Creating Actions, Avg Shot Distance, 90s Played (Total Minutes/90)
summary(PremScorersGLM2)
```

```{r}
PremScorersGLM3 <- glm(Gls ~ PrgC...23 + Sh + CPA + TO...78 + PKatt...16 + SoT90 + SoTPercent , family = "poisson", data=Prem) 
#Variables - Progressive Carries, Shots, Carries into Penalty Area, Take-Ons, Pentalty Kick Attempts, Shots on Target per 90, and Shots on Target Percentage
summary(PremScorersGLM3)
```
## Predictions for each players goals scored per season
```{r}
# Create predictions
predictions2 <- predict(PremScorersGLM2, Prem, type = 'response')
# Check the output
predictions2
```

## Combining predictions and original data into a single data frame
```{r}
# Add predictions to test data and create new data frame
predictDF <- data.frame(Prem, predictions2)
```

## Creating dataframe with only needed variables and seeing if the prediction is True (Rounded prediction = Actual goals) or False (Rounded prediction != Actual Goals)
```{r}
# Create new data frame with less columns
FinalPredictions <- select(predictDF, "Player", "Pos", "Gls", "predictions2")
# Add rounded predictions as a column
FinalPredictions$roundedPredictions <- round(FinalPredictions$predictions, 0)
# Create Boolean Column to see if real and predictions match
FinalPredictions$TFmatch <- FinalPredictions$Gls == FinalPredictions$roundedPredictions
# Check data structure 
FinalPredictions
```

##Getting the overall True/False ratio of the whole data set (NA is goalkeepers who almost never score)
```{r}
results_table <- FinalPredictions %>%
  group_by(TFmatch) %>%
  summarise(count = n())
results_table
```

## Simple linear model to get p-value for whether actual goals are a significant predictior

```{r}
fitLM <- lm(Gls ~ roundedPredictions, data = FinalPredictions)
summary(fitLM)
```

## Looking at top scorers actual goals vs. rounded predicted goals and again determining (Rounded prediction = Actual goals) or False (Rounded prediction != Actual Goals)
```{r}
TopGoalScorers <- FinalPredictions %>% 
  arrange(desc(Gls)) %>%
   select(Player, Gls, roundedPredictions, TFmatch)

TopGoalScorers
```

